---
title: "wine data"
output: html_document
---

```{r, echo=FALSE}

wineData<-read.csv('projects/nanodegree-datanalytics/project3/wineQualityReds.csv',sep=",")
library('ggplot2')
summary(wineData)
```
# Univariate
```{r, echo=FALSE}
ggplot(aes(x=quality),data=wineData) + 
  geom_histogram()
```
# Quality Histogram 
most of the wine are quality of 5 and 6
```{r, echo=FALSE}
ggplot(aes(x=sulphates),data=wineData) +
  scale_x_log10()+
  scale_x_continuous(breaks = seq(0,2, by = 0.2)) +
  geom_histogram(binwidth = .02)
```
the sulphates histogram has outliers it shows that there are some wine with more sulphate values
```{r, echo=FALSE}
ggplot(aes(x=residual.sugar),data=wineData) + 
  scale_x_continuous(breaks=seq(0,10,by=.5))+
  geom_histogram(binwidth = 1)
```
sugar value of wines lies between 1.5 and 2.5 it clearly states that most of wine have less sugary 
```{r, echo=FALSE}
ggplot(aes(x=fixed.acidity),data=wineData)+
  scale_x_continuous(breaks = seq(0,16,by=1))+
  geom_histogram()
```
acidity of wine is bit skewed left show that most of the wine are less acidity
```{r, echo=FALSE}
ggplot(aes(x=residual.sugar),data=subset(wineData,wineData$residual.sugar<8)) +
   geom_histogram()
```
most of the wine are less suagary
# Bivariate analysis
```{r, echo=FALSE}
ggplot(data=wineData,aes(x=density,y=quality)) +
  
  geom_line()
```
in the above line plot we can see that there is constant updown of desity so it does not make any sense
```{r, echo=FALSE}
#install.packages("plyr")
library(plyr)
quality_desity=ddply(wineData,~quality,summarise,desity=median(density))
ggplot(quality_desity,aes(x=desity,y=quality))+
  scale_x_continuous(breaks = seq(0.9,1,0.001))+
  geom_line() 
```
the higly disturbed desity graph is converted into sensable graph by grouping according towards quality
it also clearly depicts that high quality wine has high density value

```{r, echo=FALSE}
#install.packages('corrplot')
library(corrplot)
set.seed(1000)
corrplot(cor(wineData[,seq(2,13)]),method="number")
```
```{r, echo=FALSE}
ggplot(aes(x=citric.acid,y=fixed.acidity),data=subset(wineData,wineData$citric.acid < 0.75)) +
  geom_jitter(alpha=1/4) +
  geom_smooth()
```
the above scatter plot clearly shows that citric acid causes the acidity the outlier is also removed for the better understanding
of data

```{r, echo=FALSE}
ggplot(aes(x=pH,y=fixed.acidity),data=subset(wineData,wineData$pH>3.1))+
  geom_jitter(alpha=1/3) +
  geom_smooth()
```
the above scatter plot clearly shows that lesser ph value have more acidity
```{r, echo=FALSE}
ggplot(wineData,aes(factor(quality),alcohol)) +
  geom_boxplot()
```
the bar plot shows that high quality wine have relatively high alcohol content than lower quality wine
```{r, echo=FALSE}
ggplot(wineData,aes(factor(quality),volatile.acidity)) +
  geom_boxplot()
```
the above clearly states that low quality wine has vinegar taste but when we see quality 7 and 8 there is no significant change

```{r, echo=FALSE}
ggplot(wineData,aes(factor(quality),chlorides)) +
  geom_boxplot()
```
most of wine has same level of saltiness except 3 becuase it has a big quadrant 

```{r, echo=FALSE}
library('ggplot2')
ggplot(aes(x=total.sulfur.dioxide,y=free.sulfur.dioxide),data=wineData) + 
   geom_point()
```
here we can see that if we have more sulphur which also leads to more free sulphur but in this dataset all the point clouded in the lower point and ti shows that most of wine are less sulfur
```{r, echo=FALSE}
ggplot(aes(x=alcohol,y=density),data=wineData) +
  geom_point() +
  geom_smooth()
```
there is coorelation between density and alcohol hihger the density lower the alcohol content
```{r, echo=FALSE}
library('ggplot2')
ggplot(aes(x=total.sulfur.dioxide,y=free.sulfur.dioxide),data=wineData) + 
   geom_jitter(alpha=1/2)
```
the lowe the sulfur leads to lower the total sulfur but sulphur are clouded in bottom shows that most of the wine are less in sulfur
# multivariate analysis
```{r, echo=FALSE}
ggplot(aes(x=pH,y=volatile.acidity,color=factor(quality)),data=subset(wineData,wineData$citric.acid < 0.75)) +
  geom_point()
```
the above multivariate plot show that quality 5 wine having high acidity when compare to 7 
```{r, echo=FALSE}
wineData$citricCut=cut(wineData$citric.acid,breaks = seq(0,1,by=0.30))
ggplot(aes(x=pH,y=fixed.acidity,color=factor(citricCut)),data=subset(wineData,wineData$citric.acid < 0.75)) +
  geom_jitter(alpha=1/3)
```
we can see that there is a corelation between acidity and pH value . i  think citric acid is the reason for acidity because lower citric acid has less citric acid and high ph value
```{r, echo=FALSE}
ggplot(aes(total.sulfur.dioxide,free.sulfur.dioxide,color=citricCut),data=wineData) +
  geom_jitter(alpha=1/3) +
  facet_grid(~quality)
```
the above grid plot clearly shows that quality 5 and 6 has more sulphur and also shows that it has less
citric acid which leads to less acidiy
```{r, echo=FALSE}
ggplot(aes(x=density,y=fixed.acidity),data=wineData) +
  geom_dotplot() +
  facet_grid(~quality)
```
the wine quality 5 and 6 has high acidity and wide range of density but 3,4,7,8 has less acidity but 8 has less acidity and density
```{r, echo=FALSE}
 wineData$alcoholCut = cut(wineData$alcohol,breaks = seq(7,15,by=1.5))
  ggplot(aes(x=density,y=fixed.acidity,color=alcoholCut),data=wineData) +
  geom_point() +
  facet_grid(~quality) 
```
the above plot show that lesser density has higher alcohol content                                             
but quality 7 having lesser density abd highr alchol content





```{r}
ggplot(aes(x=total.sulfur.dioxide,y=free.sulfur.dioxide),data=wineData) + 
   geom_point()+
  facet_grid(~quality)
```
the streak of length of sulphur is less in lesser quality wine shows that lower quality wine has less in sulfur 
# Final Plot
1) ## Count of Quality of Wine
```{r, echo=FALSE}
ggplot(aes(x=quality),data=wineData) + 
  geom_histogram() +
  labs(title="                                          Count of Quality of Wine")+
  ylab("count of quality")+
  xlab("quality of wine") 
  
```
# description
the plot clearly shows that we don't have sufficent data of all quality only 5 and 6 are in high volume so taking descision we'll 
be difficult
2) ## quality on acidiy vs pH
```{r, echo=FALSE}
wineData$quality = factor(wineData$quality)
ggplot(aes(x=pH,y=volatile.acidity,color=quality),data=subset(wineData,wineData$citric.acid < 0.75)) +
  geom_point() +
  labs(title = "                                       quality on acidiy vs pH")+
  ylab("Volatile Acidity")
```
# Description 
we can see that high quality wine has less acidity but pH values is wide spreded this makes me to think the relation between ph and Volatile Acidity so volatile Acidity not affecting the pH value that much
3) ## citrus level on acidity vs pH
```{r, echo=FALSE}
wineData$citricCut=cut(wineData$citric.acid,breaks = seq(-1,2,by=0.30))
wineData$citrusLevel = factor(wineData$citricCut,labels = c("low","medium","high","very high"))
ggplot(aes(x=pH,y=fixed.acidity,color=citrusLevel),data=wineData) +
  geom_jitter(alpha=1/3) +
  labs(title="                                    citrus level on acidity vs pH")+
  ylab("Fixed Acidity")

```
# description
it clearly shows that high citrus level leads to high acidity and low pH value likewise low citrus values gives low acidity and pH value

# Reflection
- the dataset for all quality of wine is not sufficient for example 3,4,7,8 has less entries so making decision is quite bit difficult
- acidity leads to lower the ph value but volatile acidity doesn't actullay affecting the pH value of wine
- citrus level is main responsible for acidity it also affecting the pH.

